---
title: "Data Analysis Stories"
author: "Casey Canfield"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document: default
---

In this file, I will demonstrate how to perform the data analysis stories. This project uses data from:

Ludwig, J., Duncan, G. J., Gennetian, L. A., Katz, L. F., Kessler, R. C., Kling, J. R., and Sanbonmatsu, L. (2012). Neighborhood effects on the long-term well-being of low-income adults. *Science*, 337(6101), 1505–1510. 

The data is available at the [National Bureau of Economic Research website](http://www.nber.org/mtopuf/). You should download the Cell-Level PUF (Public Use Files) for the *Science* paper that was last updated on 9/21/2012.

First, I call any needed libraries.

```{r message=FALSE}


# LIBRARIES
# install.packages() if needed

library(tidyverse) # always
library(haven) # for read_dta
library(car) # for qqPlot
library(Hmisc) # for rcorr
```

Then I import data.

```{r results ='hide'}
# using results ='hide' makes it so that this doesn't have an output
# when you knit to pdf

# IMPORT DATA
mto_data <- read_dta("Data/mto_sci_puf_cells_20130206.dta") 
mto_data

# CLEAN
str(mto_data$ra_group)
mto_data$ra_group <- as_factor(mto_data$ra_group)
# ra_group needs to be a factor so that R 
# understands it's a categorical variable

# SUMMARY STATS
#summary(mto_data)
#names(mto_data) # names of all the variables in order
#objects(mto_data) # names in alphabetical order
```

Now I can perform the data analysis stories!

\pagebreak

# Data Summary Story

```{r}
# histogram showing how subjective happiness changes 
# depending on the experimental group

# identify the appropriate bin size
bw <- density(mto_data$mn_happy_scale123_z_ad, 
              kernel = "gaussian", 
              bw = "ucv")$bw

# plot histograms on top of each other
ggplot(mto_data, aes(x = mn_happy_scale123_z_ad, 
                     color = ra_group, 
                     fill=ra_group)) +
  geom_histogram(position="identity", 
                 binwidth = bw, 
                 alpha=0.5) + 
  theme_bw() +
  xlab("Experimental Group") + ylab("Subjective Happiness")
```

\pagebreak

Sometimes histograms are easier to see when they are separated. We want to understand the distributions of the data.

```{r}
# use facets to plot the histograms separately
ggplot(mto_data, aes(x=mn_happy_scale123_z_ad)) +
  geom_histogram(color="black", 
                 fill="white", 
                 binwidth = bw) +
  facet_grid(ra_group ~ .) + 
  theme_bw() +
  xlab("Experimental Group") + ylab("Subjective Happiness")
```

\pagebreak

Using box plots, we can better understand medians and outliers.

```{r}
# boxplot
ggplot(mto_data, aes(x = ra_group, y = mn_happy_scale123_z_ad)) +
  geom_boxplot() + 
  theme_bw() +
  xlab("Experimental Group") + ylab("Subjective Happiness")
```

\pagebreak

Bar plots are useful for understanding categorical variables.
```{r}
ggplot(mto_data, aes(x = ra_group)) +
  geom_bar() + 
  theme_bw() +
  xlab("Experimental Group") + ylab("Number of Observations")
```


\pagebreak

Scatter plots help us understand the relationship between two variables

```{r}
# scatter plot with regression line
ggplot(mto_data, aes(x = mn_f_ec_idx_z_ad, y = mn_happy_scale123_z_ad)) +
  geom_point() + 
  geom_smooth(method='lm') + 
  theme_bw() +
  xlab("Economic Self-Sufficiency Index") + ylab("Subjective Happiness")
```

\pagebreak

# Conditional Distribution Story

```{r}
# plot cumulative distribution
# compare data to Normal distribution

distribution_info <- mto_data %>%
  group_by(ra_group) %>%
  summarise(mean = mean(mn_happy_scale123_z_ad),
            sd = sd(mn_happy_scale123_z_ad))

set.seed(1) # Set seed for the random number generator, to reproduce results 
n.experimental <- as_tibble(rnorm(10000, 
                                  distribution_info$mean[1], 
                                  distribution_info$sd[1]))
n.section8 <- rnorm(10000, 
                    distribution_info$mean[2], 
                    distribution_info$sd[2])
n.control <- rnorm(10000, 
                   distribution_info$mean[3], 
                   distribution_info$sd[3])

# just plot experimental group data for now
mto_data %>%
  filter(ra_group == "1=Low-Poverty Voucher/Experimental") %>%
ggplot(aes(x = mn_happy_scale123_z_ad)) + 
  stat_ecdf(geom = "point") + 
  theme_bw() +
  xlab("Subjective Happiness (Experimental Group)") + ylab("Cumulative Probability") +
  stat_ecdf(aes(value), n.experimental)
```


```{r}
mto_reg <- lm(mn_happy_scale123_z_ad ~ mn_f_ec_idx_z_ad, data = mto_data)
summary(mto_reg)

# q-q plot
qqPlot(mto_reg)
```

\pagebreak

# Forecasting Story

```{r}
# cross validation

complex <- c() # Create an empty vector 
simple <- c()
set.seed(11)

for(i in 1:10){  # Loop i from 1 to 100
  train <- sample(mto_data$cell_id, 
                  2*length(mto_data$cell_id)/3, 
                  replace = FALSE)
  test <- mto_data$cell_id[ - train] 
  train1 <- lm(mn_happy_scale123_z_ad ~ mn_f_ec_idx_z_ad,
               data = mto_data[mto_data$cell_id %in% train, ]) 
  train2 <- lm(mn_happy_scale123_z_ad ~ 1,
               data = mto_data[mto_data$cell_id %in% train, ])
  test1 <- (mto_data$mn_happy_scale123_z_ad[mto_data$cell_id %in% test] -
              predict(train1, mto_data[mto_data$cell_id %in% test, ]))^2
  test2 <- (mto_data$mn_happy_scale123_z_ad[mto_data$cell_id %in% test] -
              predict(train2,mto_data[mto_data$cell_id %in% test, ]))^2
  rMSEtest1 <- sqrt(sum(test1)/length(test1)) 
  rMSEtest2 <- sqrt(sum(test2)/length(test2))
  # Append the rMSE from this iteration to vectors 
  complex <- append(complex, rMSEtest1)
  simple <- append(simple, rMSEtest2)
}

summary(complex)
summary(simple)
```

\pagebreak

# Statistical Inference Story
```{r}
# correlation
rcorr(as.matrix(mto_data[,c("mn_happy_scale123_z_ad", "mn_f_ec_idx_z_ad")]))

# hypothesis testing

# t.test
experimental <- mto_data %>%
  filter(ra_group == "1=Low-Poverty Voucher/Experimental")
control <- mto_data %>%
  filter(ra_group == "3=Control")
t.test(experimental$mn_happy_scale123_z_ad, 
       control$mn_happy_scale123_z_ad)

# regression
mto_group_lm <- lm(mn_happy_scale123_z_ad ~ ra_group, 
                    data = mto_data)
summary(mto_group_lm)

# ANOVA
mto_group_aov <- aov(mn_happy_scale123_z_ad ~ ra_group, 
                    data = mto_data) 
summary(mto_group_aov)
# post hoc test
TukeyHSD(mto_group_aov, which = 'ra_group')
```

```{r}
# resampling with bootstrap

# jacknife residuals plot
ggplot(mto_data, aes(x = fitted(mto_reg), y = rstudent(mto_reg))) +
  geom_point() + 
  geom_hline(yintercept = 0)
```

\pagebreak

# Causal Inference Story

There's no R code for this story. 
